Calculate AJCC

### 2022-12-19 Code/0_PrepareTrainingData.Rmd
Validaiton added

---
title: "Prepare Training Data"
author: "Carolien C.H.M. Maas"
date: "14-11-2022"
output: html_document
---

@@

#### Define multiple fields
```{r, eval=TRUE}
sel.train.data$multiple.fields <- ifelse(!is.na(sel.train.data$nr_SNs_Field_2)&sel.train.data$nr_SNs_Field_2>0,1,0)
```

@@

Impute recurrence time for 45 patients with missing Rec.time but Recurrence=="recurred"
```{r, eval=TRUE}
cat('Number of patients dead with recurrence without Date_1st_Rec:', sum(is.na(sel.train.data$Date_1st_Rec)&sel.train.data$Recurrence=="recurred"&sel.train.data$dead=="Yes"), '\n')
cat('Number of patients alive with recurrence without Date_1st_Rec:', sum(is.na(sel.train.data$Date_1st_Rec)&sel.train.data$Recurrence=="recurred"&sel.train.data$dead=="No"))
```

@@

Save distributions of all variables, necessary for Harrell summaries
```{r, eval=TRUE}
dd.training<-rms::datadist(training.data.to.be.imputed)
options(datadist='dd.training')
options(digits=8)
```

Relation between outcome and covariates
```{r, eval=TRUE}
training.data.to.be.imputed$logRec.time <- log(sel.train.data$Rec.time+1/365.25)
training.data.to.be.imputed$logFU.time <- log(sel.train.data$FU.time+1/365.25)
missings.table(training.data.to.be.imputed)
save(list=c("dd.training", "training.data.to.be.imputed"),file='Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata', compress=TRUE)
```

>>>>

---
title: "Prepare Training Data"
author: "Carolien C.H.M. Maas"
date: "December 2022"
output: html_document
---

@@

#### Define multiple fields
```{r, eval=TRUE}
# 11 patients for which it doesn"t hold that Tot_nr_SNs = nr_SNs_Field_1 + nr_SNs_Field_2 + nr_SNs_Field_3
sel.train.data$multiple.fields <- ifelse(!is.na(sel.train.data$nr_SNs_Field_2)&sel.train.data$nr_SNs_Field_2>0,1,0)
```

@@

Impute recurrence time for 45 patients with missing Rec.time but Recurrence=="recurred"
```{r, eval=TRUE}
cat("Number of patients dead with recurrence without Date_1st_Rec:", sum(is.na(sel.train.data$Date_1st_Rec)&sel.train.data$Recurrence=="recurred"&sel.train.data$dead=="Yes"), "\n")
cat("Number of patients alive with recurrence without Date_1st_Rec:", sum(is.na(sel.train.data$Date_1st_Rec)&sel.train.data$Recurrence=="recurred"&sel.train.data$dead=="No"))
```

@@

Save distributions of all variables, necessary for Harrell summaries
```{r, eval=TRUE}
dd.training<-rms::datadist(training.data.to.be.imputed)
options(datadist="dd.training")
options(digits=8)
```

Relation between outcome and covariates
```{r, eval=TRUE}
training.data.to.be.imputed$logRec.time <- log(sel.train.data$Rec.time+1/365.25)
training.data.to.be.imputed$logFU.time <- log(sel.train.data$FU.time+1/365.25)
missings.table(training.data.to.be.imputed)
save(list=c("dd.training", "training.data.to.be.imputed"),file="Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata", compress=TRUE)
```

### 2022-12-19 Code/0_PrepareValidationData.Rmd
Validaiton added

---
title: "Prepare Validation Data"
author: "Carolien C.H.M. Maas"
date: "14-11-2022"
output: html_document
---

@@

# load data file
original.validation.data <- foreign::read.spss(file="Z:/Project Melanoom/PaperMelanoma/Data/validation/Validation_cohort_MIA.sav",
                                               to.data.frame=TRUE, origin="01-01-1970")

# remove space after Negative # Bespreken met Robert
levels(original.validation.data$SNstatus)[levels(original.validation.data$SNstatus)=="Negative "] <- "Negative"

# all patients may be included
sel.val.data <- original.validation.data
```

@@

Merge LMM other with other
```{r, eval=TRUE}
sel.val.data$Histology[sel.val.data$Histology=="LLM"] <- "other" # Bespreken met Robert, is LLM and LMM the same?
```

#### Define multiple fields
```{r, eval=TRUE}
sel.val.data$multiple.fields <- as.numeric(!is.na(sel.val.data$Tot_nr_SNs)&sel.val.data$Tot_nr_SNs>1) # Bespreken met ask Robert
```

@@

# 339 patients have subcap with positive SN
sum(sel.val.data$SNstatus=="Positive"&sel.val.data$Dewar=="subcap"&!is.na(sel.val.data$Dewar))
```
#### Recurrence 
Define time since diagnosis in years (follow-up time)
```{r, eval=TRUE}
sel.val.data$FU.time <- (sel.val.data$Last_FU-sel.val.data$SNdate)/(24*60*60)/365.25
```

@@

Impute recurrence time for 45 patients with missing Rec.time but Recurrence=="recurred"
```{r, eval=TRUE}
cat('Number of patients dead with recurrence without Date_1st_Rec:', sum(is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Dead, melanoma       "), '\n')
cat('Number of patients alive with recurrence without Date_1st_Rec:', sum(is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Alive with Melanoma  "))
```

Set recurrence indicator
```{r, eval=TRUE}
Rec.date <- as.numeric(!is.na(sel.val.data$Date_1st_Rec)) # n=1294
Rec.date <- factor(Rec.date, labels=c("no recurrence", "recurred"))
table(sel.val.data$Status, Rec.date) # Bespreken met Robert

cat("Number of alive patients with Date_1st_Rec, but status NSR:", sum(!is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Alive NSR            "), "\n")
cat("Number of alive patients with Date_1st_Rec, but status unknown:", sum(!is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Alive, status unknown"), "\n")
cat("Number of dead patients with Date_1st_Rec, but status unknown:", sum(!is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Dead, cause unknown  "), "\n")
cat("Number of dead patients with Date_1st_Rec, but status not melanoma:", sum(!is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Dead, not melanoma   "), "\n")

# Recurrence if patient has Date_1st_Rec, Alive with melanoma, or dead with melanoma
sel.val.data$Recurrence <- as.numeric(sel.val.data$Status=="Alive with Melanoma  " | sel.val.data$Status=="Dead, melanoma       " | !is.na(sel.val.data$Date_1st_Rec)) # n=1352
sum(sel.val.data$Recurrence)
```

#### Melanoma specific mortality (MSM)
```{r, eval=TRUE}
sel.val.data$dead <- as.numeric(stringr::str_detect(sel.val.data$Status, "Dead"))

# 504 patients died due to melanoma, 102 patients died from other causes
table(sel.val.data$Status, sel.val.data$dead)

sel.val.data$MSM <- as.numeric(sel.val.data$Status=="Dead, melanoma       ")
```
#### Mitosis
Define Mitosis as Yes/No
```{r, eval=TRUE}
sel.val.data$Mitosis <- as.numeric(!is.na(sel.val.data$Mitosis))
```

@@

sel.val.data$Recurrence <- factor(sel.val.data$Recurrence, labels=c("no recurrence", "recurred"))
sel.val.data$dead <- factor(sel.val.data$dead, labels=c("No", "Yes"))
sel.val.data$MSM <- factor(sel.val.data$MSM, labels=c("Other", "Death by melanoma"))
```

# Multiple imputation
```{r, eval=TRUE}
variables.to.be.imputed <- c("Center", "SNstatus", "Sex", "Age.SN", "Ulceration",
                             "Loc_CAT", "Histology", "Breslow", "multiple.fields",
                             "Tot_SNs_neg", "Tot_SNs_pos", "Rdamcrit", "Dewar",
                             "Mitosis",
                             "Recurrence", "dead", "MSM")
validation.data.to.be.imputed <- sel.val.data[, variables.to.be.imputed]
```

Save distributions of all variables, necessary for Harrell summaries
```{r, eval=TRUE}
dd.validation<-rms::datadist(validation.data.to.be.imputed)
options(datadist='dd.validation')
options(digits=8)
```

Relation between outcome and covariates
```{r, eval=TRUE}
validation.data.to.be.imputed$logRec.time <- log(sel.val.data$Rec.time+1/365.25)
validation.data.to.be.imputed$logFU.time <- log(sel.val.data$FU.time+1/365.25)
missings.table(validation.data.to.be.imputed)
save(list=c("dd.validation", "validation.data.to.be.imputed"),file='Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.Rdata', compress=TRUE)
```

>>>>

---
title: "Prepare Validation Data"
author: "Carolien C.H.M. Maas"
date: "December 2022"
output: html_document
---

@@

# load data file
original.validation.data <- foreign::read.spss(file="Z:/Project Melanoom/PaperMelanoma/Data/validation/Validation_cohort_MIA.sav",
                                               to.data.frame=TRUE, origin="01-01-1970")
cat("Number of patients in data set:", nrow(original.validation.data), "\n")

# remove space after Negative
levels(original.validation.data$SNstatus)[levels(original.validation.data$SNstatus)=="Negative "] <- "Negative"

# exclude 4 patients without Status (don"t know if they alive or dead)
cat("Number of patients without Status:", sum(original.validation.data$Status=="Unknown              "), "\n")
sel.val.data <- original.validation.data[original.validation.data$Status!="Dead, MCC            ",]
# exculde 1 patient with cause of death MCC
sel.val.data <- sel.val.data[sel.val.data$Status!="Unknown              ",]
cat("Number of included patients:", nrow(sel.val.data))
```

@@

Merge LMM other with other
```{r, eval=TRUE}
sel.val.data$Histology[sel.val.data$Histology=="LLM"] <- "other" # LLM and LMM is the same
```

#### Define multiple fields
```{r, eval=TRUE}
sel.val.data$multiple.fields <- as.numeric(!is.na(sel.val.data$Tot_nr_SNs)&sel.val.data$Tot_nr_SNs>1)
```

@@

# 339 patients have subcap with positive SN
sum(sel.val.data$SNstatus=="Positive"&sel.val.data$Dewar=="subcap"&!is.na(sel.val.data$Dewar))
```
#### Recurrence 
Set recurrence indicator
```{r, eval=TRUE}
# Recurrence if patient has Date_1st_Rec, Alive with melanoma, or dead with melanoma
sel.val.data$Recurrence <- as.numeric(!is.na(sel.val.data$Date_1st_Rec)) # n=1352
# sel.val.data$Recurrence <- as.numeric(sel.val.data$Status=="Alive with Melanoma  " | sel.val.data$Status=="Dead, melanoma       " | !is.na(sel.val.data$Date_1st_Rec)) # n=1352
sel.val.data$Recurrence <- factor(sel.val.data$Recurrence, labels=c("no recurrence", "recurred"))
```

Define time since diagnosis in years (follow-up time)
```{r, eval=TRUE}
sel.val.data$FU.time <- (sel.val.data$Last_FU-sel.val.data$SNdate)/(24*60*60)/365.25
# median: 4.1629021
```

@@

Set recurrence time equal to follow-up time for those without recurrence
```{r, eval=TRUE}
sel.val.data$Rec.time[sel.val.data$Recurrence=="no recurrence"|is.na(sel.val.data$SNdate)]<-sel.val.data$FU.time[sel.val.data$Recurrence=="no recurrence"|is.na(sel.val.data$SNdate)]
table(sel.val.data$Status, sel.val.data$Recurrence)
```

#### Melanoma specific mortality (MSM)
```{r, eval=TRUE}
sel.val.data$dead <- as.numeric(stringr::str_detect(sel.val.data$Status, "Dead"))
sel.val.data$dead <- factor(sel.val.data$dead, labels=c("No", "Yes"))
table(sel.val.data$Status, sel.val.data$dead)
sel.val.data$MSM <- as.numeric(sel.val.data$Status=="Dead, melanoma       ")
```
#### Mitosis
Define Mitosis as Yes/No
```{r, eval=TRUE}
sel.val.data$Mitosis <- as.numeric(!is.na(sel.val.data$Mitosis))
```

@@

sel.val.data$MSM <- factor(sel.val.data$MSM, labels=c("Other", "Death by melanoma"))
```

# Multiple imputation
```{r, eval=TRUE}
variables.to.be.imputed <- c("Center", "SNstatus", "Sex", "Age.SN", "Ulceration",
                             "Loc_CAT", "Histology", "Breslow", "multiple.fields",
                             "Tot_SNs_neg", "Tot_SNs_pos", "Rdamcrit", "Dewar",
                             "Mitosis",
                             "Recurrence", "dead", "MSM", "Rec.time", "FU.time")
validation.data.to.be.imputed <- sel.val.data[, variables.to.be.imputed]
```

Save distributions of all variables, necessary for Harrell summaries
```{r, eval=TRUE}
dd.validation<-rms::datadist(validation.data.to.be.imputed)
options(datadist="dd.validation")
options(digits=8)
```

Relation between outcome and covariates
```{r, eval=TRUE}
validation.data.to.be.imputed$logRec.time <- log(sel.val.data$Rec.time+1/365.25)
validation.data.to.be.imputed$logFU.time <- log(sel.val.data$FU.time+1/365.25)
missings.table(validation.data.to.be.imputed)
save(list=c("dd.validation", "validation.data.to.be.imputed"),file="Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.Rdata", compress=TRUE)
```

### 2022-12-19 Code/10_AJCC.Rmd
Validaiton added

---
title: "Calculate AJCC"
author: "Carolien C.H.M. Maas"
date: "14-11-2022"
output: html_document
---

@@

# load libraries
# remotes::install_github("CHMMaas/PredictionTools")
library(PredictionTools)  # to plot calibration
```
# Obtain C-index of AJCC7 and AJCC8
```{r, eval=TRUE}
calculate.AJCC.7.8 <- function(imputed.data=NULL, type=NULL, S.Rec.5=NULL, S.MSM.5=NULL){
  ####
  #### AJCC 7
  ####
  TNM.7 <- c()
  AJCC.7 <- c()
  cindex.AJCC.7 <- rep(0, m)
  cindex.AJCC.7.se <- rep(0, m)
  cindex.AJCC.7.pos <- rep(0, m)
  cindex.AJCC.7.se.pos <- rep(0, m)
  cindex.AJCC.7.neg <- rep(0, m)
  cindex.AJCC.7.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.7.i <- rep("", nrow(imputation.i))
    AJCC.7.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

@@

    # compute C-index for only negative sentinel node patients
    rc.neg <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Negative"], S.Rec.5[imputation.i$SNstatus=="Negative",])
    cindex.AJCC.7.neg[i]<-rc.neg["C Index"]
    cindex.AJCC.7.se.neg[i]<-rc.neg["S.D."]/2
  }
  # combine C-index for each imputed data set
  cindex.mi.AJCC.7<-Rubin.combine(cindex.AJCC.7,cindex.AJCC.7.se)
  cindex.mi.AJCC.7.pos<-Rubin.combine(cindex.AJCC.7.pos,cindex.AJCC.7.se.pos)
  cindex.mi.AJCC.7.neg<-Rubin.combine(cindex.AJCC.7.neg,cindex.AJCC.7.se.neg)

  # save C-index and confidence interval
  utils::write.table(rbind(c("All patients:", sprintf('%.2f', cindex.mi.AJCC.7$est),
                paste0("[", sprintf('%.2f', cindex.mi.AJCC.7$est+qnorm(.025)*cindex.mi.AJCC.7$se),
                        "; ", sprintf('%.2f', cindex.mi.AJCC.7$est+qnorm(.975)*cindex.mi.AJCC.7$se), "]")),
              c("Positive SN patients:", sprintf('%.2f', cindex.mi.AJCC.7.pos$est),
                    paste0("[", sprintf('%.2f', cindex.mi.AJCC.7.pos$est+qnorm(.025)*cindex.mi.AJCC.7.pos$se),
                           "; ", sprintf('%.2f', cindex.mi.AJCC.7.pos$est+qnorm(.975)*cindex.mi.AJCC.7.pos$se), "]")),
              c("Negative SN patients:", sprintf('%.2f', cindex.mi.AJCC.7.neg$est),
                    paste0("[", sprintf('%.2f', cindex.mi.AJCC.7.neg$est+qnorm(.025)*cindex.mi.AJCC.7.neg$se),
                           "; ", sprintf('%.2f', cindex.mi.AJCC.7.neg$est+qnorm(.975)*cindex.mi.AJCC.7.neg$se), "]"))),
              file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/AJCC.7.Cindex.', type, '.txt'), 
              row.names=FALSE, col.names=FALSE, sep=",")

@@

  # save C-index and confidence interval
  utils::write.table(rbind(c("All patients:", sprintf('%.2f', cindex.mi.AJCC.8$est),
                          paste0("[", sprintf('%.2f', cindex.mi.AJCC.8$est+qnorm(.025)*cindex.mi.AJCC.8$se),
                                 "; ", sprintf('%.2f', cindex.mi.AJCC.8$est+qnorm(.975)*cindex.mi.AJCC.8$se), "]")),
                    c("Positive SN patients:", sprintf('%.2f', cindex.mi.AJCC.8.pos$est),
                          paste0("[", sprintf('%.2f', cindex.mi.AJCC.8.pos$est+qnorm(.025)*cindex.mi.AJCC.8.pos$se),
                                 "; ", sprintf('%.2f', cindex.mi.AJCC.8.pos$est+qnorm(.975)*cindex.mi.AJCC.8.pos$se), "]")),
                    c("Negative SN patients", sprintf('%.2f', cindex.mi.AJCC.8.neg$est),
                          paste0("[", sprintf('%.2f', cindex.mi.AJCC.8.neg$est+qnorm(.025)*cindex.mi.AJCC.8.neg$se),
                                 "; ", sprintf('%.2f', cindex.mi.AJCC.8.neg$est+qnorm(.975)*cindex.mi.AJCC.8.neg$se), "]"))),
              file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/AJCC.8.Cindex.', type, '.txt'), 
              row.names=FALSE, col.names=FALSE, sep=",")

@@

  if (type=="training"){
    table.AJCC.7.8 <- cbind(c('1.1', '1.2', '2.1', '2.2', '2.3', '3.1', '3.2', '3.3', '3.4'),
                          c(round(rowMeans(table.AJCC.7), 0), 0),
                          round(rowMeans(table.AJCC.8)))
  } else if (type=="validation"){
    table.AJCC.7.8 <- cbind(c('1.1', '1.2', '2.1', '2.2', '2.3', '3.1', '3.2', '3.3', '3.4'),
                          c(round(rowMeans(table.AJCC.7), 0), 0),
                          c(0, round(rowMeans(table.AJCC.8), 0)))
  }
  utils::write.table(table.AJCC.7.8,
              file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/table.AJCC.7.and.8.', type, '.txt'),
              row.names=FALSE, col.names=FALSE, sep=",")
}
load('Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData')
calculate.AJCC.7.8(imputed.data=imputed.training.data, S.Rec.5=S.Rec.5.training, S.MSM.5=S.MSM.5.training, type='training')
load('Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData')
calculate.AJCC.7.8(imputed.data=imputed.validation.data, S.Rec.5=S.Rec.5.validation, S.MSM.5=S.MSM.5.validation, type='validation')
```

>>>>

---
title: "Calculate AJCC"
author: "Carolien C.H.M. Maas"
date: "December 2022"
output: html_document
---

@@

# load libraries
library(PredictionTools)  # to plot calibration
library(compareC)         # compare AJCC to model
```
# Obtain C-index of AJCC7 and AJCC8
```{r, eval=TRUE}
calculate.AJCC.7.8 <- function(imputed.data=NULL, type=NULL, S.Rec.5=NULL){
  ####
  #### AJCC 7
  ####
  TNM.7 <- c()
  AJCC.7 <- c()
  cindex.AJCC.7 <- rep(0, m)
  cindex.AJCC.7.se <- rep(0, m)
  cindex.AJCC.7.pos <- rep(0, m)
  cindex.AJCC.7.se.pos <- rep(0, m)
  cindex.AJCC.7.neg <- rep(0, m)
  cindex.AJCC.7.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.7.i <- rep("", nrow(imputation.i))
    AJCC.7.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

@@

    # compute C-index for only negative sentinel node patients
    rc.neg <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Negative"], S.Rec.5[imputation.i$SNstatus=="Negative",])
    cindex.AJCC.7.neg[i]<-rc.neg["C Index"]
    cindex.AJCC.7.se.neg[i]<-rc.neg["S.D."]/2
  }

  # combine C-index for each imputed data set
  cindex.mi.AJCC.7<-Rubin.combine(cindex.AJCC.7,cindex.AJCC.7.se)
  cindex.mi.AJCC.7.pos<-Rubin.combine(cindex.AJCC.7.pos,cindex.AJCC.7.se.pos)
  cindex.mi.AJCC.7.neg<-Rubin.combine(cindex.AJCC.7.neg,cindex.AJCC.7.se.neg)

  # save C-index and confidence interval
  AJCC.7.table <- rbind(c("All patients:", sprintf("%.2f", cindex.mi.AJCC.7$est),
                paste0("[", sprintf("%.2f", cindex.mi.AJCC.7$est+qnorm(.025)*cindex.mi.AJCC.7$se),
                        "; ", sprintf("%.2f", cindex.mi.AJCC.7$est+qnorm(.975)*cindex.mi.AJCC.7$se), "]")),
              c("Negative SN patients:", sprintf("%.2f", cindex.mi.AJCC.7.neg$est),
                    paste0("[", sprintf("%.2f", cindex.mi.AJCC.7.neg$est+qnorm(.025)*cindex.mi.AJCC.7.neg$se),
                           "; ", sprintf("%.2f", cindex.mi.AJCC.7.neg$est+qnorm(.975)*cindex.mi.AJCC.7.neg$se), "]")),
              c("Positive SN patients:", sprintf("%.2f", cindex.mi.AJCC.7.pos$est),
                    paste0("[", sprintf("%.2f", cindex.mi.AJCC.7.pos$est+qnorm(.025)*cindex.mi.AJCC.7.pos$se),
                           "; ", sprintf("%.2f", cindex.mi.AJCC.7.pos$est+qnorm(.975)*cindex.mi.AJCC.7.pos$se), "]")))

@@

  # save C-index and confidence interval
  AJCC.8.table <- rbind(c("All patients:", sprintf("%.2f", cindex.mi.AJCC.8$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8$est+qnorm(.025)*cindex.mi.AJCC.8$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8$est+qnorm(.975)*cindex.mi.AJCC.8$se), "]")),
                    c("Negative SN patients", sprintf("%.2f", cindex.mi.AJCC.8.neg$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8.neg$est+qnorm(.025)*cindex.mi.AJCC.8.neg$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8.neg$est+qnorm(.975)*cindex.mi.AJCC.8.neg$se), "]")),
                    c("Positive SN patients:", sprintf("%.2f", cindex.mi.AJCC.8.pos$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8.pos$est+qnorm(.025)*cindex.mi.AJCC.8.pos$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8.pos$est+qnorm(.975)*cindex.mi.AJCC.8.pos$se), "]")))

@@

  if (type=="training"){
    AJCC.7.8.count <- cbind(c("1.1", "1.2", "2.1", "2.2", "2.3", "3.1", "3.2", "3.3", "3.4"),
                          c(round(rowMeans(table.AJCC.7), 0), 0),
                          round(rowMeans(table.AJCC.8)))
  } else if (type=="validation"){
    AJCC.7.8.count <- cbind(c("1.1", "1.2", "2.1", "2.2", "2.3", "3.1", "3.2", "3.3", "3.4"),
                          c(round(rowMeans(table.AJCC.7), 0), 0),
                          c(0, round(rowMeans(table.AJCC.8), 0)))
  }
  return(list(AJCC.7=AJCC.7, AJCC.7.table=AJCC.7.table,
              AJCC.8=AJCC.8, AJCC.8.table=AJCC.8.table, AJCC.7.8.count=AJCC.7.8.count))
}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
train.AJCC <- calculate.AJCC.7.8(imputed.data=imputed.training.data, S.Rec.5=S.Rec.5.training, type="training")
load("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData")
validation.AJCC <- calculate.AJCC.7.8(imputed.data=imputed.validation.data, S.Rec.5=S.Rec.5.validation, type="validation")
```

# Compare C-index of AJCC and developed model
```{r, eval=TRUE}
for (edition in c(7, 8)){
  for (patient.type in c("All", "Negative", "Positive")){
    model.C <- c()
    se.C <- c()
    diff <- c()
    se <- c()
    for (i in 1:m){
      # compute linear predictor
      lp.i <- predict(f.mi.BS.Rec.5,
                    newdata=mice::complete(imputed.validation.data, i),
                    type="lp")

      if (patient.type=="All"){
        lp.i <- lp.i
        S <- S.Rec.5.validation
        AJCC <- eval(parse(text=paste0("validation.AJCC$AJCC.", edition, "[, i]")))
      } else if (patient.type=="Positive"){
        lp.i <- lp.i[validation.data.to.be.imputed$SNstatus=="Positive"]
        S <- S.Rec.5.validation[validation.data.to.be.imputed$SNstatus=="Positive",]
        AJCC <- eval(parse(text=paste0("validation.AJCC$AJCC.", edition, "[validation.data.to.be.imputed$SNstatus==\"Positive\", i]")))
      } else if (patient.type=="Negative"){
        lp.i <- lp.i[validation.data.to.be.imputed$SNstatus=="Negative"]
        S <- S.Rec.5.validation[validation.data.to.be.imputed$SNstatus=="Negative",]
        AJCC <- eval(parse(text=paste0("validation.AJCC$AJCC.", edition, "[validation.data.to.be.imputed$SNstatus==\"Negative\", i]")))
      }

      # compute predictions from model
      p.i <- 1-exp(-h0.Rec*exp(lp.i))

      # compare AJCC to model
      compare.i <- compareC::compareC(S[,1], S[,2], -p.i, -AJCC)

      model.C <- c(model.C, compare.i$est.c["Cxy"])
      se.C <- c(se.C, sqrt(compare.i$est.varCxy))

      # save differences and SE of difference
      diff <- c(diff, compare.i$est.diff_c)
      se <- c(se, sqrt(compare.i$est.vardiff_c))
    }
    assign(paste0("diff.", patient.type, ".", edition), Rubin.combine(diff, se))
    assign(paste0("model.", patient.type), Rubin.combine(model.C, se.C))
  }
}

# save in Table
compare.AJCC.model <- rbind(c("All patients:", 
                sprintf("%.2f", diff.All.7$est),
                paste0("[", sprintf("%.2f", diff.All.7$est+qnorm(.025)*diff.All.7$se),
                        "; ", sprintf("%.2f", diff.All.7$est+qnorm(.975)*diff.All.7$se), "]"),
                sprintf("%.2f", diff.All.8$est),
                paste0("[", sprintf("%.2f", diff.All.8$est+qnorm(.025)*diff.All.8$se),
                        "; ", sprintf("%.2f", diff.All.8$est+qnorm(.975)*diff.All.8$se), "]")),
              c("Negative SN patients:", 
                sprintf("%.2f", diff.Negative.7$est),
                paste0("[", sprintf("%.2f", diff.Negative.7$est+qnorm(.025)*diff.Negative.7$se),
                       "; ", sprintf("%.2f", diff.Negative.7$est+qnorm(.975)*diff.Negative.7$se), "]"),
                sprintf("%.2f", diff.Negative.8$est),
                paste0("[", sprintf("%.2f", diff.Negative.8$est+qnorm(.025)*diff.Negative.8$se),
                       "; ", sprintf("%.2f", diff.Negative.8$est+qnorm(.975)*diff.Negative.8$se), "]")),
              c("Positive SN patients:", 
                sprintf("%.2f", diff.Positive.7$est),
                paste0("[", sprintf("%.2f", diff.Positive.7$est+qnorm(.025)*diff.Positive.7$se),
                       "; ", sprintf("%.2f", diff.Positive.7$est+qnorm(.975)*diff.Positive.7$se), "]"),
                sprintf("%.2f", diff.Positive.8$est),
                paste0("[", sprintf("%.2f", diff.Positive.8$est+qnorm(.025)*diff.Positive.8$se),
                       "; ", sprintf("%.2f", diff.Positive.8$est+qnorm(.975)*diff.Positive.8$se), "]")))

CIndex.val <- rbind(c("All patients:",
                sprintf("%.2f", model.All$est),
                paste0("[", sprintf("%.2f", model.All$est+qnorm(.025)*model.All$se), 
                      "; ", sprintf("%.2f", model.All$est+qnorm(.975)*model.All$se), "]")),
                c("Negative SN patients:",
                sprintf("%.2f", model.Negative$est), 
                paste0("[", sprintf("%.2f", model.Negative$est+qnorm(.025)*model.Negative$se), 
                      "; ", sprintf("%.2f", model.Negative$est+qnorm(.975)*model.Negative$se), "]")),
                c("Positve SN patients:",
                sprintf("%.2f", model.Positive$est), 
                paste0("[", sprintf("%.2f", model.Positive$est+qnorm(.025)*model.Positive$se), 
                      "; ", sprintf("%.2f", model.Positive$est+qnorm(.975)*model.Positive$se), "]")))
```

# Write to Excel
```{r, eval=TRUE}
utils::write.table(cbind(train.AJCC$AJCC.7.table, rep("", 3), train.AJCC$AJCC.8.table[,-1]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.train.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
utils::write.table(cbind(CIndex.val, rep("", 3),
                   validation.AJCC$AJCC.7.table[, -1], rep("", 3), 
                   compare.AJCC.model[, 2:3], rep("", 3),
                   validation.AJCC$AJCC.8.table[,-1], rep("", 3),
                   compare.AJCC.model[, 4:5]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.validation.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
utils::write.table(cbind(train.AJCC$AJCC.7.8.count, rep("", 3), validation.AJCC$AJCC.7.8.count[,-1]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.7.8.count.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
```
